KidsAIR: merge health datasets

### 2020-04-07 child_health_data_cleaning.Rmd
LRTI tracking updates

# Load and merge files from different locations
```{r}
wmt_medicalvisit_initial <- read_xlsx("Input/WMT/wmt_medicalvisit_20200227.xlsx") %>% 
  replace_with_na_all(condition = ~.x == -999) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Notes)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, med_visit_date) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

nn_medicalvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200227.xlsx") %>% 
  replace_with_na_all(condition = ~.x == -999) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, med_visit_date) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

ak_medicalvisit_initial <- read_xlsx("Input/AK/ak_medicalvisit_20200227.xlsx") %>% 
  replace_with_na_all(condition = ~.x == -999) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "AK",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, med_visit_date) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

>>>>

# Load and merge files from different locations
```{r}
wmt_medicalvisit_initial <- read_xlsx("Input/WMT/wmt_medicalvisit_20200227.xlsx") %>% 
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

nn_medicalvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200227.xlsx") %>%
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

ak_medicalvisit_initial <- read_xlsx("Input/AK/ak_medicalvisit_20200227.xlsx") %>% 
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "AK",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

### 2020-04-07 lrti_tracking.Rmd
LRTI tracking updates

---
title: "Kids LRTI tracking"
author: "Ethan Walker"
date: "Started 6 April 2020, Updated 6 April 2020"
output: pdf_document 
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

@@

```{r}
lrti_data <- health_exposure_data_sampling_day %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         med_visit_date:med_visit_notes, chw_dr_relook:chw_ari_comments,
         csq_date:csq_comments, med_rec_visit_date:med_rec_comments) %>% 
  filter(!is.na(visit_date)) %>% 
  filter(area == "WMT") %>% 
  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  distinct(visit_date, .keep_all = TRUE) %>% 
  mutate(trash_date = ymd("2099-01-01"),
         still_sick_date = if_else(csq_still_sick == "Yes", csq_date, trash_date),
         med_rec_ari_date = if_else(med_rec_ari != "No ARI", med_rec_visit_date, trash_date),
         med_visit_ari_date = if_else(med_visit_ari != "No", med_visit_date, trash_date)) %>% 
  pivot_longer(cols = c("still_sick_date", "med_rec_ari_date", "med_visit_ari_date"),
               names_to = "lrti_visits", values_to = "lrti_dates") %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  mutate(lrti_start_date_new = if_else(!is.na(csq_first_symptom_date),
                                       csq_first_symptom_date, lrti_dates)) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, lrti_dates) %>% 
  distinct(visit_date, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num) %>% 
  distinct(lrti_start_date_new, .keep_all = TRUE)

table(lrti_data$med_visit_ari)
```

>>>>

---
title: "Kids LRTI tracking"
author: "Ethan Walker"
date: "Started 6 April 2020, Updated 7 April 2020"
output: pdf_document 
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
```

@@

# Diagnosing LRTI from the data

Is there a specific protocol Paul uses for diagnosing LRTI, or is it case-by-case
based on the given information for each child? We'll need to have some kind of
protocol/algorithm for publications, I would think.

Other questions:
- How far has the team made it (date/winter) in diagnosing LRTI? Winter 17/18?
- Has this process been done/started for NN/AK, or just WMT?

```{r}
# LRTI Diagnosis data
lrti_diagnosis_data <- health_exposure_data_sampling_day %>% 
  # select relevant health data
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         med_visit_date:med_visit_mean_temp,
         csq_date:csq_comments, 
         med_rec_visit_date, med_rec_dr_diagnosis, med_rec_ari, med_rec_comments) %>% 
  # Currently only have this data for WMT - need to get new files and updates dataset
  filter(area == "WMT") %>% 
  # This is the most broad way of distinguishing if a child was sick
  # Need to get Paul's protocol for diagnosis to filter down from here
  filter(csq_sick_2_weeks == "Yes")
```

The filtering below narrows down the LRTI definition too much; we are missing cases
where children had LRTI but did not visit a healthcare facility and have a med rec.
If we only use these cases, it could bias the results.

There are instances of LRTI in our records that weren’t confirmed in the medical record. 
There are other instances where the medical record diagnosed LRTI, but our records 
overturned it. What is our criteria for diagnosing LRTI, where is it documented, 
and how can we be sure it is consistent in diagnosing LRTI?

```{r}
# Further filtering and re-ordering of variables
lrti_diagnosis_data_filtered <- lrti_diagnosis_data %>% 
  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
  # If a home CSQ was done on same visit date as initial phone CSQ, filter for only the home CSQ
  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
         csq_first_symptom_date:csq_comments) %>% 
  # Were they referred (med visit) or did they see a Dr (CSQ)?
  filter(med_visit_referred == "Yes" | csq_see_dr == "Yes") %>% 
  # If yes, did we get medical records?
  filter(!is.na(med_rec_visit_date)) %>% 
  # If yes, was there an ARI diagnosis?
  filter(med_rec_ari == "ARI" | med_rec_ari == "Possible ARI") %>% 
  group_by(child_id_num) %>% 
  distinct(med_rec_visit_date, .keep_all = TRUE)
```

# Defining start date, end date, and at-risk period for each diagnosed LRTI case

First, I need to know (or make) exactly which variable is a final diagnosed case 
of LRTI in the data. Then, I can create a method in R to identify start/end/at-risk.
The code below is a start at this, using Nick's handoff documents/algorithm.

```{r}
# LRTI start date, end date, and at-risk period identification
lrti_data <- health_exposure_data_sampling_day %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         med_visit_date:med_visit_notes, chw_dr_relook:chw_ari_comments,
         csq_date:csq_comments, med_rec_visit_date:med_rec_comments) %>% 
  filter(area == "WMT") %>% 
  filter(csq_sick_2_weeks == "Yes") %>% 
  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
         csq_first_symptom_date:csq_comments)
  mutate(trash_date = ymd("2099-01-01"),
         still_sick_date = if_else(csq_still_sick == "Yes", csq_date, trash_date),
         med_rec_ari_date = if_else(med_rec_ari != "No ARI", med_rec_visit_date, trash_date),
         med_visit_ari_date = if_else(med_visit_ari != "No", med_visit_date, trash_date)) %>% 
  pivot_longer(cols = c("still_sick_date", "med_rec_ari_date", "med_visit_ari_date"),
               names_to = "lrti_visits", values_to = "lrti_dates") %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  mutate(lrti_start_date_new = if_else(!is.na(csq_first_symptom_date),
                                       csq_first_symptom_date, lrti_dates)) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date,
         csq_sick_2_weeks, csq_still_sick, med_rec_ari, med_visit_ari) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, lrti_dates) %>% 
  distinct(visit_date, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num) %>% 
  distinct(lrti_start_date_new, .keep_all = TRUE)

table(lrti_data$med_visit_ari)
table(health_exposure_data_sampling_day$csq_sick_2_weeks)
```

### 2020-04-07 merge_health_datasets.Rmd
LRTI tracking updates

---
title: "KidsAIR: merge health datasets"
author: "Ethan Walker"
date: "Started 27 March 2020, Updated 6 April 2020"
output: pdf_document
---

@@

health_merged1 <- kids_csq %>% 
  select(-home_id) %>% 
  mutate(visit_date = csq_date) %>% 
  full_join(kids_med_visit, by = c("area", "home", "home_winter_id", "treatment",
                                   "winter_id", "child_id_num", "visit_date")) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         treatment, csq_date:med_visit_notes) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

@@

# Merge full health data with exposure data
```{r}
exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")

exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")

exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")

sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")

>>>>

---
title: "KidsAIR: merge health datasets"
author: "Ethan Walker"
date: "Started 27 March 2020, Updated 7 April 2020"
output: pdf_document
---

@@

health_merged1 <- kids_csq %>% 
  select(-home_id) %>% 
  mutate(visit_date = csq_date) %>% 
  full_join(kids_med_visit, by = c("area", "home", "home_winter_id", "treatment",
                                   "winter_id", "child_id_num", "visit_date")) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         treatment, csq_date:med_visit_mean_temp) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

@@

# Merge full health data with exposure data
```{r}
exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
